---
title: "PA1_template"
author: "AT"
date: "6 octobre 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Week 2 Peer-graded Assignment  

##Introduction
This Markdown document develops the coding requested in the peer-graded assigment for the course Reproducible Research.   

## Data
For this peer-graded assigment, the data to use is based on the measurements of different monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up as  part of the movement "quantified self".  
This data, from those personal activity monitoring devices, collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.  
The data is available is available in the current git folder an was download it from this link: (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)  

The variables included in this dataset are:  

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA  
* date: The date on which the measurement was taken in YYYY-MM-DD format  
* interval: Identifier for the 5-minute interval in which measurement was taken  

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Code  
  
### Loading and preprocessing the data

0.1 Check if the extra libraries are installed (some peers don't have zip package)
```{r 0.1}
x <- c("zip" )
new<- x[!(x %in% installed.packages()[,"Package"])]
if(length(new)) install.packages(new)
rm(x,new)
```

0.2 Read libraries
```{r 0.2}
library(zip)
library(lattice)
library(ggplot2)
library(dplyr)
```

0.3 Check if data set is already downloaded in the current working directory and download it if is not available.
```{r 0.3}
if (file.exists("activity.csv")){print("Already exists")
} else {
      file1 <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(file1, destfile = "x.zip", method = "curl")
      unzip("x.zip") 
}
```

0.4 Read information
```{r 0.4}
act <- read.csv("activity.csv")
```
### What is mean total number of steps taken per day?

1.0 Obtain total number of steps taken per day (with out NA)
```{r 1.0}
act_clean <-  act[complete.cases(act), ]
steps<- tapply(act_clean$steps, act_clean$date, FUN=sum)

```

2.0 Create histogram with total number of steps taken each day
```{r 2.0}
histogram(~ steps, data = act_clean,
          main = "Histogram wo NA",
          xlab = "Steps",
          ylab = "Frecuency",
          col = "chocolate3")
dev.copy(png, file = "plot1.png", width=480, height=480)
dev.off()
```

3.0 Mean and median total number of steps taken per day
```{r 3.0}
steps<- tapply(act_clean$steps, act_clean$date, FUN=sum)
mean_steps<- mean(steps, na.rm = T)
mean_steps
median_steps<- median(steps, na.rm = T)
median_steps
```
The results show mean =10766.19 steps and median = 10765 steps.

### What is the average daily activity pattern?

4.0 Create plot with average steps by period of 5 minutes interval
```{r 4.0}
steps5min <- as.table(tapply(act_clean$steps, act_clean$interval, mean, na.rm = TRUE))
plot(as.numeric(names(steps5min)),steps5min,
     xlab = "Interval", ylab = "Steps",
     main = "Average Daily Activity Pattern", type = "l")
```

5.0 Obtain maximum number of steps 5-min period
```{r 5.0}
m <- max(steps5min)
maxsteps <- steps5min[steps5min==m]
maxsteps[]
```
The results shows that in average, 08:35 hrs is the time with more steps

###Imputing missing values

6.0 Total number of rows with NAs
```{r 6.0}
missingNA <- sum(!complete.cases(act))
missingNA
```
The total number of rows with NA values are 2304 5-min periods. This represents 8 hours of missing data,or 13.11% of  the total data.

7.0 Strategy for filling the missing values
```{r 7.0}
f_replace<- function(x){ replace(x, is.na(x), mean(x, na.rm = TRUE))}
```

8.0 New dataset with the NA values replaced by the mean
```{r 8.0}
new_data <- act %>%
      group_by(interval) %>%
      mutate(steps = f_replace(steps))
```

9.0 Reviewing if there is any NA value 
```{r 9.0}
missingNA <- sum(!complete.cases(new_data))
missingNA
```
It is confirmed that all NA values were replaced

10.0 Create a new dataset that  equal to the original dataset but with the missing data filled in.
```{r 10.0}
steps2<- tapply(new_data$steps, new_data$date, FUN=sum)
```

11.0 Histogram total number of steps taken each day (NA filled)
```{r 11.0}
histogram(~ steps, data = new_data,
          main = "Histogram with NA",
          xlab = "Steps",
          ylab = "Frecuency",
          col = "blue1")
dev.copy(png, file = "plot2.png", width=480, height=480)
dev.off()
```

12.0 Calculate and report the mean/median total number of steps taken per day.
```{r 12.0}
steps<- tapply(new_data$steps, new_data$date, FUN=sum)
n_mean_steps<- mean(steps, na.rm = T)
n_mean_steps
n_median_steps<- median(steps, na.rm = T)
n_median_steps
```

13.0 What is the impact NA on the estimates of the total daily number of steps?

```{r 13.0}
Difference_mean= n_mean_steps - mean_steps
Difference_mean
Difference_median = median_steps- n_median_steps
Difference_median
```
The difference in mean steps is 0 and the difference in median steps is 1.18, thus is minimal.

### Are there differences in activity patterns between weekdays and weekends?

14.0 Create a new factor variable in the dataset with two levels “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r 14.0}
new_data$wd <- weekdays( as.Date(new_data$date))
new_data$wd <- ifelse (new_data$wd=="Saturday" | new_data$wd=="Sunday", "Weekday", "Weekend" )
new_data2 <- aggregate(new_data$steps, by = list(new_data$wd, new_data$interval), mean)
colnames (new_data2) <-(c("day", "interval", "number_steps"))
```
15.0 Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r 15.0}
xyplot(number_steps ~ interval| day, data = new_data2, layout = c(1, 2), type= "l")
dev.copy(png, file = "plot3.png", width=480, height=480)
dev.off()
```
End of code
